---
title: "DANL 200-02: Beer Market Data Analysis"
subtitle: "Final Project"
author: 
  - Jayden Hewitt
  - Syed Zahoor Raza
  - Andre Bianchi
  - Eric Smith

toc: true
toc-title: "Table of Contents"
toc-depth: 2
number-sections: true

fig-width: 9

execute:
  code-tools: true
  message: false
  warning: false

from: markdown+emoji
---
```{r setup}
#| include: false
library(tidyverse)
library(skimr)
library(ggthemes)
library(hrbrthemes)

theme_set(theme_ipsum()+
          theme(strip.background =element_rect(fill="lightgray"),
                axis.title.x = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(10,0,0,0)),
                axis.title.y = 
                  element_text(angle = 0,
                               size = rel(1.33),
                               margin = margin(0,10,0,0))
                )
          )
```

```{r}
beer_data <- read_csv("https://bcdanl.github.io/data/beer_markets_all.csv")
```

How Do the Prices Per Fluid Ounce Vary from One
Brand to Another?
Analyzing price per fluid ounce across brands is crucial for both consumers and businesses within
the beer industry. Customers are always looking to get the most out of their purchases, and
comparing the cost per fluid ounce of various beer brands helps them identify which brands are the
best deals on beer. This is crucial for consumers on a tight budget who might value quantity above
brand awareness. Thus, they can make informed purchasing decisions using this information.
Moreover, the analysis can also help with market segmentation. By understanding how price impacts
various consumer groups, businesses can tailor marketing and promotions. For instance, to draw
customers, a retailer may decide to price some goods cheaply or to give premium brands larger
margins. If data indicates that a particular beer brand has a higher price point than rivals, a grocery
store chain may choose to launch a promotion on it to draw in more consumers or boost sales
volume. In public policy, regulatory authorities may also find this data useful to evaluate possible
price gouging or to ensure fair competition.
Below, I use the R code to calculate the average price per fluid ounce for each brand of beer and
standard deviation to see how much the price per fluid ounce varies within each brand. The R code
to calculate these is as follows:
```{r}
summary_by_brand <- beer_data %>%
group_by(brand) %>%
summarize(mean_price_floz = mean(price_per_floz),
sd_price_floz = sd(price_per_floz))
```

